---
title: "Landscapemetrics Exercise"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T, message = F, warning = F)
library(tidyverse)
library(landscapemetrics)
library(landscapetools)
library(raster)
library(knitr)
library(kableExtra)
```

## Landscape metrics calculation exercise using the [Landscapemetrics](https://github.com/r-spatialecology/landscapemetrics) library

### Download raster data:

[Enhanced Vegetation Index](https://en.wikipedia.org/wiki/Enhanced_vegetation_index) downloaded from [Nasa Giovanni](https://giovanni.gsfc.nasa.gov/giovanni/giovanni-login-redirect?code=9778b03b2b4906b69a84134c82d0d2eba9f2ffef0c5f94ce46a0ba002e5a30ed&state=https%3A%2F%2Fgiovanni%2Egsfc%2Enasa%2Egov%2Fgiovanni%2F) 

```{r}
raster<-raster("./raster/GIOVANNI-g4.timeAvgMap.MYD13C2_006_CMG_0_05_Deg_Monthly_EVI.20200101-20200531.74W_6N_72W_7N.tif")

check_landscape(raster) %>% 
  kable(caption="EVI features") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T,
                full_width = F, 
                position = "left")
```
The raster has a geographic CRS, is necesary to transform to projected CRS using [epsg:3117](https://spatialreference.org/ref/epsg/magna-sirgas-colombia-bogota-zone/) 
```{r}
crs(raster)<- "+init=epsg:3117 +proj=longlat +datum=WGS84 +no_defs"

check_landscape(raster) %>% kable()
```

#### Summary of raster data

```{r ,echo = F, fig.show="hide" }
summary(raster)
f <- hist(raster, main="Raster histogram", col="steelblue", breaks=20)

dat <- data.frame(counts= f$counts,breaks = f$mids)
```

#### Raster histogram 

```{r}
ggplot(dat, aes(x = breaks, y = counts, fill =counts)) + ## Note the new aes fill here
  geom_bar(stat = "identity",alpha = 1)+
  labs(x="Pixel value",y="Frequency",
       title = "Raster histogram",
       caption = "EVI from Nasa Giovanny:\nGIOVANNI-g4.timeAvgMap.MYD13C2_006_CMG_0_05_Deg_Monthly_EVI.20200101-20200531.74W_6N_72W_7N")+
  scale_x_continuous(breaks = seq(-0.1,1,0.1),
                   labels = seq(-0.1,1,0.1))+
  scale_fill_gradient(low="grey80", high="white")+          ## to play with colors limits
  theme_light()+
  theme(plot.title=element_text(size=19),
        plot.caption = element_text(size = 7, color="grey60"),
        panel.background = element_rect(fill="steelblue"),
        panel.grid.major = element_line(size = 0.3, linetype = 'dotted',
                                colour = "white"),
        panel.grid.minor = element_line(size = 0.3, linetype = 'dotted',
                                colour = "white"),
        legend.position = "none")

```

### Raster reclasification

#### clasification matrix

This is where the different classes and the cutoff points of each one are defined. In this example we are going to reclassify the raster into tree classes.

```{r}
# create classification matrix
reclass_df <- c(0.181, 0.3, 1,
              0.3, 0.4, 1,
             0.4, 0.5, 2,
             0.5, Inf, 3)
reclass_df
# reshape the object into a matrix with columns and rows
reclass_m <- matrix(reclass_df,
                ncol = 3,
                byrow = TRUE)
reclass_m
```

And here is where the reclassification process is made

```{r}
# reclassify the raster using the reclass object - reclass_m
raster_classified <- reclassify(raster,reclass_m)
```
```{r,echo = F, fig.show="hide"}
g <- hist(raster_classified, main="Raster histogram", col="steelblue", breaks=3)

dat2 <- data.frame(counts= g$counts,breaks = g$mids)
```

### Reclassification result 

```{r}
ggplot(dat2, aes(x = breaks, y = counts, fill =counts)) + ## Note the new aes fill here
  geom_bar(stat = "identity",alpha = 1)+
  labs(x="Class",y="",
       title = "Number of pixels in each class",
       caption = "EVI from Nasa Giovanny:\nGIOVANNI-g4.timeAvgMap.MYD13C2_006_CMG_0_05_Deg_Monthly_EVI.20200101-20200531.74W_6N_72W_7N")+
  scale_x_continuous(breaks = seq(1,3,1),
                   labels = seq(1,3,1))+
  scale_fill_gradient(low="white", high="white")+          ## to play with colors limits
  theme_light()+
  theme(plot.title=element_text(size=19),
        plot.caption = element_text(size = 7, color="grey60"),
        panel.background = element_rect(fill="steelblue"),
        panel.grid.major = element_line(size = 0.3, linetype = 'dotted',
                                colour = "white"),
        panel.grid.minor = element_line(size = 0.3, linetype = 'dotted',
                                colour = "white"),
        legend.position = "none")


```

### Plot the reclassified raster

```{r}

# plot reclassified data
show_landscape(raster_classified)


check_landscape(raster_classified)
```

Here we see the tree different patch

### landscape metrics  

- Table of metrics by each patch

```{r}
list_lsm(level = 'patch') 

metrics<- calculate_lsm(raster_classified, level = "patch") 

kable(metrics, caption = "Landscape Metrics by patch") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T,
                full_width = F, 
                position = "left") %>% 
  scroll_box(width = "400px", height = "400px")
```

- Table of metrics by patch class

```{r}
metrics %>% 
  group_by(class, metric) %>% 
  summarise(Val=sum(value)) %>% 
  kable(caption="Landscape Metrics by patch class") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T,
                full_width = F, 
                position = "left") %>% 
  scroll_box(width = "250px", height = "400px")
```


### Patches plot

```{r}

show_patches(raster_classified, labels = T)

# show patches of all classes
show_patches(raster_classified, class = "all", labels = FALSE)
```

### Patches area

```{r}
# fill patch according to area
show_lsm(raster_classified, what = "lsm_p_area", class = "global", label_lsm = TRUE)
```

### Correlation matrix

```{r}
show_correlation(metrics)
```

